---
title: "Homework 8"
author: "Jess Wikle"
date: "3/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data manipulations using the `dplyr` package

###### This homework assignment focus on data manipulation in R. Complete these problems using the `dplyr` package, and if you have time, try each of these exercises using the subsetting methods in base R that have been covered in lecture.

#### 1. Examine the structure of the iris dataset. How many observations and variables are in the dataset?

```{r}
library(dplyr)
glimpse(iris)
```

The `iris` dataset contains 150 observations of 5 variables. Species is a factor variable and the plant attributes (length and width of sepal and petal) are double, or numeric variables.


#### 2. Create a new data frame `iris1` that contains only the species virginica and versicolor with sepal lengths longer than 6 cm and sepal widths longer than 2.5 cm. How many observations and variables are in the dataset?

```{r}
iris1 <- filter(iris, Species %in% c("virginica", "versicolor"), 
                Sepal.Length > 6, 
                Sepal.Width > 2.5)
glimpse(iris1)
```

The dataframe `iris1` contains 56 observations of the same 5 variables as above. 



#### 3. Now, create a `iris2` data frame from `iris1` that contains only the columns for Species, Sepal.Length, and Sepal.Width. How many observations and variables are in the dataset?

```{r}
iris2 <- select(iris1, Species, Sepal.Length, Sepal.Width)
head(iris2)
glimpse(iris2)
```

The data frame `iris2` contains 56 observations of 3 variables.




#### 4.Create an `iris3` data frame from `iris2` that orders the observations from largest to smallest sepal length. Show the first 6 rows of this dataset.
```{r}
iris3 <- arrange(iris2, desc(Sepal.Length))
head(iris3)
```


#### 5. Create an `iris4` data frame from `iris3` that creates a column with a sepal area (length * width) value for each observation. How many observations and variables are in the dataset?

```{r}
iris4 <- mutate(iris3, Sepal.Area = Sepal.Length * Sepal.Width)
glimpse(iris4)
```

The data frame `iris4` has 56 observations of 4 variables.




#### 6. Create `iris5` that calculates the average sepal length, the average sepal width, and the sample size of the entire `iris4` data frame and print `iris5`.
```{r}
iris5 <- summarize(iris4, mean_sepal_length = round(mean(Sepal.Length),2),
                   mean_sepal_width = round(mean(Sepal.Width),2),
                   sample_size = n())
print(iris5)
```


#### 7. Finally, create `iris6` that calculates the average sepal length, the average sepal width, and the sample size for each species of in the `iris4` data frame and print `iris6`.
```{r}
iris6 <- iris4 %>% 
  group_by(Species) %>% 
  summarize(mean_sepal_length = round(mean(Sepal.Length),2),
                   mean_sepal_width = round(mean(Sepal.Width),2),
                   sample_size = n())
print(iris6)
```


#### 8. In these exercises, you have successively modified different versions of the data frame `iris1`, `iris2`, `iris3`, `iris4`, `iris5`, `iris6`. At each stage, the output data frame from one operation serves as the input for the next.

#### A more efficient way to do this is to use the pipe operator `%>%`from the `tidyr` package. See if you can rework all of your previous statements into an extended piping operation that uses `iris` as the input and generates `iris6` as the output.
```{r}
iris6 <- iris %>% 
  filter(Species %in% c("virginica", "versicolor"), 
                Sepal.Length > 6, 
                Sepal.Width > 2.5) %>% 
  select(Species, Sepal.Length, Sepal.Width) %>% 
  arrange(desc(Sepal.Length)) %>% 
  mutate(Sepal.Area = Sepal.Length * Sepal.Width) %>% 
  group_by(Species) %>% 
  summarize(mean_sepal_length = round(mean(Sepal.Length),2),
                   mean_sepal_width = round(mean(Sepal.Width),2),
                   sample_size = n())
print(iris6)
  

```

