---
title: "Homework 5"
author: "Jess Wikle"
date: "2/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Working with matrices, lists, and data frames

### 1. 

Assign to the variable n_dims a single random integer between 3 and 10
```{r}
n_dims <- sample(3:10, 1)


```

Create a vector of consecutive integers from 1 to n_dims2
```{r}
vec <- 1:n_dims^2
```


Use the sample function to randomly reshuffle these values
```{r}
vec1 <- sample(vec)
```


Create a square matrix with these elements
```{r}
mat <- matrix(vec1, nrow = n_dims)
```


Print out the matrix
```{r}
print(mat)
```


Find a function in r to transpose the matrix
```{r}
mat1 <- t(mat)
```


Print it out again and note how it has changed
```{r}
print(mat1)
```


Calculate the sum and the mean of the elements in the first row and the last row
```{r}
sum(mat1[1,])
mean(mat1[1,])

sum(mat1[n_dims,])
mean(mat1[n_dims,])
```


Read about the eigen() function and use it on your matrix
```{r}
eig <- eigen(mat1)
print(eig)
```


Look carefully at the elements of $values and $vectors. What kind of numbers are these?
Dig in with the typeof() function to figure out their type.

```{r}
typeof(eig$values)
typeof(eig$values[1])
typeof(eig$vectors)
typeof(eig$vectors)[1]
```

If have set your code up properly, you should be able to re-run it and create a matrix of different size because n_dims will change.

### 2.
Create a list with the following named elements:
my_matrix, which is a 4 x 4 matrix filled with random uniform values
my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
my_letters, which is a 26-element vector of all the lower-case letters in random order.

```{r}
my_list <- list(my_matrix = matrix(runif(4), nrow = 2), 
                my_logical = runif(100) > 0.35,
                my_letters = sample(letters))
print(my_list)
str(my_list)
```
Create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.

```{r}
my_list2 <- list(my_list$my_matrix[2,2], 
                 my_list$my_logical[2], 
                 my_list$my_letters[2])
print(my_list2)
```

Use the typeof() function to confirm the underlying data types of each component in this list
```{r}
typeof(my_list2[[1]])
typeof(my_list2[[2]])
typeof(my_list2[[3]])
```

Combine the underlying elements from the new list into a single atomic vector with the c() function.
```{r}
vec2 <- c(my_list2[[1]], my_list2[[2]], my_list2[[3]])
print(vec2)
```

What is the data type of this vector?
```{r}
typeof(vec2)
```

Create a data frame with two variables (= columns) and 26 cases (= rows).
Call the first variable my_unis and fill it with a random uniform value from 0 to 10
Call the second variable my_letters and fill it with 26 capital letters in random order.

```{r}
my_df <- data.frame()[26, 2]
my_df$my_unis <- runif(1:26, min = 0, max = 10)
my_df$my_letters <- sample(LETTERS)
my_df <- as.data.frame(my_df)
print(my_df)
```

For the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.
```{r}
my_df[sample(1:26,4),1] <- NA
print(my_df)
```

For the first variable, write a single line of R code to identify which rows have the missing values.
```{r}
which(is.na(my_df$my_unis))
```


For the second variable, sort it in alphabetical order
```{r}

my_df <- my_df[order(my_df$my_letters),]
print(my_df)
```


Calculate the column mean for the first variable.
```{r}
mean(my_df$my_unis, na.rm = TRUE)
```

