---
title: "Homework 12"
author: "Jess Wikle"
date: "4/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# packages
library(ggplot2)
library(ggthemes)
library(wesanderson)
library(patchwork)
```

## Advanced ggplotting

For this exerise, use your newly-developed ggplot chops to create some nice graphs from your own data (If you do not have a good data frame to use for graphics, use one of the many built-in data frames from R (other than mpg, which we are using in class)). Experiment with different themes, theme base sizes, aesthetics, mappings, and faceting. When you are finished, try exporting them to high quality pdfs, jpgs, eps files, or other formats that you would use for submission to a journal.

#### Categorical data exploration
```{r}
library(ggplot2)
library(ggthemes)
library(wesanderson)

data <- read.csv("cwmforplot.csv")
head(data)

# I want both of these columns to be in specific, non-alphabetical order
data$year2 <- factor(data$year, levels = c("preharvest", "postharvest"))
data$treat2 <- factor(data$treat, levels = c("Control", "Resistance", "Resilience", "Transition"))

# set background
theme_set(theme_classic())

# boxplot of volume of coarse woody material in four treatments, pre and post-treatment, grayscale(ish)

p1 <- ggplot(data, aes(x = treat2, y = m3ha, fill = year2)) +
  geom_boxplot() +
  scale_fill_manual(values = c("ivory3", "ivory4")) +
  ylim(0, 150) + # make y axis reasonable size, even though it hides major outliers
  labs(title = "Volume of Downed Woody Material in Four Silvicultural Treatments",
       subtitle = "greyscale boxplot",
       x = "Treatment",
       y = "Volume (cubic meters per hectare)") +
  theme(legend.title = element_blank())
print(p1)

# with color
p1 <- ggplot(data, aes(x = treat2, y = m3ha, fill = year2)) +
  geom_boxplot() +
  scale_fill_manual(values = wes_palette("GrandBudapest1", 2)) +
  ylim(0, 150) +
   labs(title = "Volume of Downed Woody Material in Four Silvicultural Treatments",
       subtitle = "color boxplot",
       x = "Treatment",
       y = "Volume (cubic meters per hectare)") +
  theme(legend.title = element_blank())
print(p1)

# with color and points
p1 <- ggplot(data, aes(x = treat2, y = m3ha, 
                       fill = year2, color = year2)) +
  geom_boxplot() +
  scale_fill_manual(values = wes_palette("GrandBudapest1", 2)) +
  geom_point(position = position_jitter(width = 0.1, height = 0.7),
              size = 2, alpha = 0.6) +
  scale_color_manual(values = c("goldenrod", "firebrick")) +
  ylim(0, 150) +
   labs(title = "Volume of Downed Woody Material in Four Silvicultural Treatments",
       subtitle = "boxplot with color and points",
       x = "Treatment",
       y = "Volume (cubic meters per hectare)") +
  theme(legend.title = element_blank())
print(p1)

# Edward Tufte-style boxplot (minimal and aesthetically pleasing)
# haven't figured out how to do include split for pre-post treatment

theme_set(theme_tufte())
p1 <- ggplot(data, aes(x = treat2, y = m3ha)) +
  geom_tufteboxplot() +
  ylim(0, 150) +
   labs(title = "Volume of Downed Woody Material in Four Silvicultural Treatments",
       subtitle = "tufte boxplot",
       x = "Treatment",
       y = "Volume (cubic meters per hectare)") +
  theme(legend.title = element_blank())
print(p1)
  
# can facet to make up for inability to split
p1 <- ggplot(data, aes(x = year2, y = m3ha)) +
  geom_tufteboxplot() +
  ylim(0, 150) +
   labs(title = "Volume of Downed Woody Material in Four Silvicultural Treatments",
       subtitle = "tufte boxplot",
       x = "Time",
       y = "Volume (cubic meters per hectare)") +
  facet_grid(.~treat2)
print(p1)

```

#### Continuous data exploration
```{r}
data <- read.csv("dsiforplot1.csv")
head(data)


# Disturbance severity index (DSI) is predictor variable
# Response is proportion of regeneration that consists of 
# sugar maple(shade tolerant, less disurbance tolerant)
# yellow birch (thrives in disturbance)

p1 <- ggplot(data, aes(x = DSI, y = proportion, color = species))
p1 + geom_smooth(method = 'lm')

# minus error bands
p1 + geom_smooth(method = 'lm', se = FALSE)

# plus economist theme
p1 + geom_smooth(method = 'lm', se = FALSE) +
  labs(title = "Proportion of Regeneration that is Sugar Maple and Yellow Birch\n as Predicted by Disturbance Severity") +
  theme_economist_white()

# plus better color scheme
p1 + geom_smooth(method = 'lm', se = FALSE) +
  labs(title = "Proportion of Regeneration that is Sugar Maple and Yellow Birch\n as Predicted by Disturbance Severity") +
  scale_color_economist() +
  theme_economist()

# plus wsj theme 
p1 + geom_smooth(method = 'lm', se = FALSE) +
  scale_color_wsj() +
  labs(title = "Proportion of Regeneration that is Sugar Maple and Yellow Birch\n as Predicted by Disturbance Severity") +
  theme_wsj()
# (odd that it removes axis labels...). I guess that's the theme though
p1 + geom_smooth(method = 'lm', se = FALSE) +
  labs(title = "Proportion of Regeneration that is Sugar Maple and Yellow Birch\n as Predicted by Disturbance Severity") +
  labs(x = "DSI", y = "proportion") +
  scale_color_wsj() +
  theme_wsj()

# Tufte theme
p1 + geom_smooth(method = 'lm', se = FALSE) +
  scale_color_manual(values = c("grey30", "grey60")) +
  labs(title = "Proportion of Regeneration that is Sugar Maple and Yellow Birch\n as Predicted by Disturbance Severity") +
  theme_tufte()

# solarized theme
p1 + geom_smooth(method = 'lm', se = FALSE) +
  labs(title = "Proportion of Regeneration that is Sugar Maple and Yellow Birch\n as Predicted by Disturbance Severity") +
  scale_color_solarized() +
  theme_solarized()
```

#### Non-theme changes
```{r}
data <- read.csv("/Users/jessicawikle/Desktop/R/ASCC/biomasstreatment.csv", header = T)

for(i in 1:nrow(data)){
  if(data$year[i] == "2017"){
    data$timing[i] <- "preharvest"
  }else data$timing[i] <- "postharvest"
}
data$timefac <- factor(data$timing, levels = c("preharvest", "postharvest"))
data$compfac <- factor(data$Component, levels = c("CWM", "Snags", "Stumps", "LiveTrees"))
data$treatmentfac <- factor(data$Treatment, levels = c("Control", "Resistance", "Resilience", "Transition"))

cols <- c("LiveTrees" = 'darkgreen',
          "Stumps" = 'goldenrod1',
          "Snags" = 'coral4',
          "CWM" = 'dodgerblue')

p1 <- ggplot(data, aes(x = timefac, y = Biomass))

p1 + geom_bar(stat = 'summary', 
              fun.y = 'sum',
             colour = 'black', 
             size = 1, 
             width = 0.7) +
  geom_bar(stat = 'identity', 
              aes(fill = compfac),
           width = 0.7)+
  labs(x = "Time",
       y = "Biomass (Mg/ha)",
       fill = "Component") +
  facet_grid(~treatmentfac) + 
  scale_fill_manual(values = cols) +
  theme(aspect.ratio = 3) +
  theme_classic()
```

#### Exporting figures
```{r}
data <- read.csv("dsiforplot1.csv")
head(data)

p1 <- ggplot(data, aes(x = DSI, y = proportion, color = species)) + geom_smooth(method = 'lm', se = FALSE) +
  scale_color_manual(values = c("grey30", "grey60")) +
  labs(title = "Proportion of Regeneration that is Sugar Maple and Yellow Birch\n as Predicted by Disturbance Severity") +
  theme_tufte()

ggsave(filename = "testplot.pdf", plot = p1, device = "pdf",
       width = 11, height = 8.5, units = "in")

ggsave(filename = "testplot.jpg", plot = p1, device = "jpg",
       width = 6, height = 4, units = "in")

 
```


